### 2015-01-26 sharp-sigma-one-plus.tex
26 ene

\def\A{{\frak A}}
\def\C{{\cal C}}
\def\F{{\cal F}}
\def\L{{\cal L}}
\def\N{\mathbb{N}}
\def\P{\vec{P}}
\def\R{\vec{R}}
\def\S{\vec{S}}
\def\X{\vec{X}}
\def\Y{\vec{Y}}
\def\Z{\vec{Z}}
%% a - arity of \X / arity of assignments \P to \X
%% b - arity of predicates in \S
%% c - arity of auxiliar predicates/variables
\def\d{\vec{d}} %% counted elements
\def\e{\vec{e}} %% counted elements
%% f - counting function
%% g - other functions
%% h - other functions
%% i - index
%% j - index
%% k - emergency index / size of tuple
%% l - emergency index / size of tuple
\def\l{\vec{\ell}}
%% m - size of variable tuple
%% n - size of predicate tuple
%% o - not used
%% p - 
%% q - 
%% r - size of \X / \P
\def\s{\vec{s}}
%% t - size of \S
\def\u{\vec{u}} %% auxiliary variables
\def\v{\vec{v}} %% auxiliary variables
\def\w{\vec{w}} %% auxiliary variables
\def\x{\vec{x}} %% quantified variables
\def\y{\vec{y}} %% auxiliary variables
\def\z{\vec{z}} %% open variables
\def\ep{\vec{o}}
\def\ga{\vec{p}}

@@

We define the vocabulary $\L = \{ S_1,\dots,S_t, \leq \}$, where $S_1,\dots,S_t$ have arity $b_1,\dots,b_t$. Let
\begin{eqnarray*}
\Truc &=& \{\A \mid \A \text{ is an } \L \text{-structure with a finite domain } A \text{ such that} \\
&& \leq \text{ is interpreted as a total order for } A \}.
\end{eqnarray*}
We also define a set of second order variables ${\cal X} = \{ X_i \mid i\in\N \}$ where $X_i$ has arity $a_i$, and for every $n \in \N$ there are infinite variables in ${\cal X}$ of arity $n$. A quantifier-free $\L$-formula is defined by the following grammar:
\begin{eqnarray*}
\varphi &::=& x = y \ \mid \ S_i(x_1,\dots,x_{b_i}), i \in \{1,\dots,t\} \ \mid \ x \leq y \ \mid \\
&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \\ 
&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi),
\end{eqnarray*}
where $x,y$ and $x_i$ are first order variables for every $i$. We now define an extended quantifier-free $\L$-formula as follows:
\begin{eqnarray*}
\varphi &::=& \alpha, \alpha \text{ is an FO-formula over } \L  \ \mid \\
&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \ \\
&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi).
\end{eqnarray*}
Let $\Y = (Y_1,\dots,Y_q)$ be a tuple of second-order variables of arity $c_1,\ldots,c_q$, and let $\y$ be a tuple of first order variables. For every $\L$-formula $\psi(\Y,\y)$, we define the function $f_{\psi(\Y,\y)}:\Truc \to \N$ as follows:
\begin{eqnarray*}
f_{\psi(\Y,\y)}(\A) &=& \mid \{ \langle \P, \e \rangle \mid \A \models \psi(\P,\e) \} \mid,
\end{eqnarray*}
for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_q)$ is a tuple of predicates of arity $a_1,\ldots,a_q$, and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,q\}$, and $\e$ is a tuple of elements in $\A$.

Let $\X = (X_1,\dots,X_r)$, and let $\z$ be a tuple of variables. A function $f:\Truc \to \N$ is in $\E1$ if there exists an extended quantifier-free $\L$-formula $\varphi(\x,\X,\z)$ such that $f = f_{\exists \x \: \varphi(\x,\X,\z)}.$\\

@@

For a given function $f$, we define $f \dotminus 1$ as follows:
\begin{eqnarray*}
f \dotminus 1(\A) =
\begin{cases}
f(\A)-1, & \text{if }f(\A) > 0 \\
0, & \text{if }f(\A) = 0.
\end{cases}
\end{eqnarray*}
for every structure $\A$. A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$, it holds that $f \dotminus 1 \in \F$.

@@

\begin{theo}
$\E1$ is closed under substraction by one.
\end{theo}
\begin{proof}
For every $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$, let $\z = (z_1,\dots,z_d)$ and let $\X = (X_1,\dots,X_r)$. There are three possibilities regarding the size of the tuples of free variables $\z$ and $\X$: (1) $d>0$ and $r=0$ (2) $d=0$ and $r>0$ (3) $d,r>0$. This separates the proof in three cases:
\begin{enumerate}

@@

\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,\dots,x_d)$ and $\X = (X_1,\dots,X_r)$. That is,
\begin{eqnarray*}
f(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \exists \x \ \varphi(\x,\P) \} \mid \label{f1},
\end{eqnarray*}
for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_r)$ and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$. For the time being, suppose that
\begin{eqnarray}
\varphi(\x,\X) &=& \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)
\end{eqnarray}
where $n$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, $\y$ is a $p$-tuple of variables in $\x$, $\varphi^{-}(\X,\y)$ is a conjunction of negated predicates in $\X$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\beta(\x)$ is an FO-formula over $\L$. We also assume that $(\x_1,\dots,\x_n,\y) = \x$. As an example, the following formula is of this form:
\begin{align*}
\varphi(\x,\X) =  X_1(x_1,x_2) \wedge \neg X_1(x_3,x_4) \wedge (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 ) \wedge \forall z\big( S_1(x_1,z) \big),
\end{align*}
where $\x = (x_1,x_2,x_3,x_4)$ and $\X = (X_1)$. Here, $n = 1$, $\lambda(1) = \lambda(2) = 1$, $\x_1 = (x_1,x_2)$ and $\y = (x_3,x_4)$. Moreover, $\varphi^{-}(\X,\y) = \neg X_1(x_3,x_4)$, $\theta(\x) = (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4)$, which defines a total order on the partition of $\x$ $\{\{x_1,x_3,x_4\},\{x_2\}\}$, and $\beta(\x) = \forall z\big( S_1(x_1,z) \big)$.

@@

\varphi^\prime(\x,\X) = \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)\wedge\exists\u_1\cdots\exists\u_n\bigg[\alpha_{\min}(\u_1,\dots,\u_n) \wedge \\ \bigg(\bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\u_i) \bigg) \vee \bigvee_{i=1}^r \exists \v\Big( X_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big) \bigg) \bigg] \label{f2}.

@@

\begin{lemma}  \label{lemmaone}

@@

\item[(a)] By contradiction, suppose that $\A\not\models\exists\x\,\varphi(\x,\P)$. That is, there is no assignment $\s$ to $\x$ such that $\A\models\varphi(\s,\P).$ Since $\d$ is such that $\A\models \bigwedge_{i=1}^n P_{\lambda(i)}(\d_i)$ and $\A\models \alpha(\d)$, it follows that $\A\not\models\varphi^{-}(\P,\ep)$ (since $\alpha(\d)=\theta(\d)\wedge\beta(\d)$). Therefore, there is an $i\in\{1,\ldots,n\}$ such that $\neg P_{\lambda(i)}(\d_i)$ appears in $\varphi^{-}(\P,\ep)$. Then, there is an $a_{\lambda(i)}$-tuple $\z$ in $\y$ such that $\neg X_{\lambda(i)}(\z)$ appears in $\varphi^{-}(\X,\x)$. We know that either (1) $\theta(\x)\models \z = \x_i$, (2) $\theta(\x)\models \z < \x_i$ or (3) $\theta(\x)\models \z > \x_i$. Considering that (2) and (3) are not possible since $\A\models\theta(\d)$, and both $\z$ and $\x_i$ are assigned the value $\d_i$, then $\theta(\x)\models \z = \x_i$. But if this is the case, then $X_{\lambda(i)}(\x_i), \neg X_{\lambda(i)}(\z)$ and $\z = \x_i$ are all logical consequences of $\varphi(\x,\X)$ from which we conclude that $\A\not\models\exists\x\,\varphi(\x,\R^\prime)$ for every possible assignment $\R^\prime$ to $\X$, which leads to a contradiction.
\item[(b)] By the construction of $\P$, we see that 
$$\A\not\models\bigvee_{i = 1}^{n}\neg P_{\lambda(i)}(\d_i) \text{ and that } \A\not\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big).$$ However, the only possible assignment to $\alpha_{\min}(\x_1,\dots,\x_n)$ is $(\d_1,\dots,\d_n)$. Then, $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$.
\item[(c)] By contradiction, let $\P^\prime \neq \P$ be such that $\A\models\exists\x\,\varphi(\x,\P^\prime)$ and $\A\not\models\exists\x\,\varphi^\prime(\x,\P^\prime)$. We consider two cases: first, suppose that $\P^\prime$ is missing a tuple of $\P$. Let $i\in\{1,\ldots,n\}$ such that in $\d_i$ is not in $\P^\prime$, then $\A\models \neg P^\prime_i(\d_i)$. And so,$$\A\models \bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\d_i) \bigg), $$ from which we conclude that $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$. Second, suppose there is some predicate $\P^\prime_i$ in $\P^\prime$ which has a tuple that $\P_i$ does not have. If this is the case, then $$\A\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big),$$ so $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$. On both cases, we have a contradiction.

@@

Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\R)$. Let $\P$ be an arbitrary assignment to $\X$. Since $\A\not\models\exists\x\,\varphi(\x,\P)$, we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\psi(\x,\P))$ for any formula $\psi(\x,\P)$. It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime(\x,\R)$. And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1.$
\end{proof}

We now continue with the general case, in which $\varphi(\x,\X)$ is an arbirtary extended quantifier-free $\L$-formula. By using a standard DNF transformation algorithm and considering FO-formulas over $\L$ as literals, we can find formulas $\hat{\varphi}_i(\x,\X)$ with $i\in\{1,\ldots,\hat{k}\}$ such that
\begin{eqnarray*}
\varphi(\x,\X) &\equiv& \hat{\varphi}_1(\x,\X) \vee \hat{\varphi}_2(\x,\X) \vee \dots  \vee \hat{\varphi}_{\hat{k}}(\x,\X),
\end{eqnarray*}
where, for every $i\in\{1,\ldots,\hat{k}\}$, 
\begin{eqnarray*}
\hat{\varphi}_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \hat{\varphi}^{-}_i(\X,\y_i)  \wedge \beta_i(\x).
\end{eqnarray*}
The function $\lambda_i$ is defined analogously to the first part of the proof, the tuple $\x_{i,j}$ has $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, $\y_i$ has $p_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x.$ The formulas $\hat{\varphi}^{-}_i(\X,\y_i)$, $\beta_i(\x)$, are also defined analogously. Let $g$ be a function that counts the number of possible orders over partitions on a $d$-tuple of variables. Let the formulas $\theta^i(\x)$ for $i\in\{1,\ldots,g(d)\}$ represent each of these orders over $\x$. Note that for every formula $\psi(\x)$,
\begin{eqnarray*}
\exists\x\:\psi(\x) &\equiv& \exists\x(\psi(\x)\wedge\theta^1(\x)) \vee \cdots \vee \exists\x(\psi(\x)\wedge\theta^{g(d)}(\x)).
\end{eqnarray*}
We define the following formulas $\check{\varphi}_i(\x,\X)$, for $i\in\{1,\ldots,\check{k}\},$ as follows: 
\begin{eqnarray*}
\check{\varphi}_1(\x,\X) &=& \hat{\varphi}_1(\x,\X) \wedge \theta^1(\x), \\
& \vdots & \\
\check{\varphi}_{g(d)}(\x,\X) &=& \hat{\varphi}_1(\x,\X) \wedge \theta^{g(d)}(\x), \\
\check{\varphi}_{g(d)+1}(\x,\X) &=& \hat{\varphi}_2(\x,\X) \wedge \theta^1(\x), \\
& \vdots & \\
\check{\varphi}_{2\cdot g(d)}(\x,\X) &=& \hat{\varphi}_2(\x,\X) \wedge \theta^{g(d)}(\x), \\
& \vdots & \\
\check{\varphi}_{(\hat{k}-1) g(d)+1}(\x,\X) &=& \hat{\varphi}_{\hat{k}}(\x,\X) \wedge \theta^1(\x), \\
& \vdots & \\
\check{\varphi}_{\hat{k} \cdot g(d)}(\x,\X) &=& \hat{\varphi}_{\hat{k}}(\x,\X) \wedge \theta^{g(d)}(\x).
\end{eqnarray*}
Note that $\check{k} = \hat{k} \cdot g(d).$ Having every disjunct with a total order allows us to eliminate the ones that are unsatisfiable for every $\L-$structure $\A$, that is, every $i\in\{1,\ldots,\check{k}\}$ such that for every assignment $\s$ to $\x,$ $f_{\check{\varphi}_i(\s,\X)} = 0.$ Let $k$ be the number of disjuncts that are left after eliminating the unsatisfiable disjuncts. We order the remaining disjuncts and define the disjuncts $\varphi_i(\x,\X)$ such that for every $j\in\{1,\ldots,\check{k}\}$ where $\check{\varphi}_j(\x,\X)$ is satisfiable, there is an $i\in\{1,\ldots,k\}$ such that $\varphi_i(\x,\X) = \check{\varphi}_j(\x,\X).$

As seen in lemma \ref{lemmaone}, for every $\L-$structure $\A$ and $i\in\{1,\ldots,k\}$, if there is an assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$ but $f_{\varphi_i(\s,\X)}(\A) = 0,$ then $f_{\exists\x\:\varphi_i(\x,\X)}(\A) = 0$, so it follows that for every assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$, then $f_{\exists\x\:\varphi_i(\x,\X)}(\A) > 0.$

>>>>

\def\A{{\frak A}}
\def\C{{\cal C}}
\def\F{{\cal F}}
\def\L{{\cal L}}
\def\N{\mathbb{N}}
\def\P{\vec{P}}
\def\R{\vec{R}}
\def\S{\vec{S}}
\def\X{\vec{X}}
\def\Y{\vec{Y}}
\def\Z{\vec{Z}}
%% a - arity of \X / arity of assignments \P to \X
%% b - arity of predicates in \S
%% c - arity of auxiliar predicates/variables
\def\c{\vec{c}} %% super auxiliar elements
\def\d{\vec{d}} %% counted elements
\def\e{\vec{e}} %% counted elements
%% f - counting function
%% g - other functions
%% h - other functions
%% i - index
%% j - index
%% k - emergency index / size of tuple
%% l - emergency index / size of tuple
\def\l{\vec{\ell}}
%% m - size of variable tuple
%% n - size of predicate tuple
%% o - not used
%% p - 
%% q - 
%% r - size of \X / \P
\def\s{\vec{s}}
%% t - size of \S
\def\t{\vec{t}}
\def\u{\vec{u}} %% auxiliary variables
\def\v{\vec{v}} %% auxiliary variables
\def\w{\vec{w}} %% auxiliary variables
\def\x{\vec{x}} %% quantified variables
\def\y{\vec{y}} %% auxiliary variables
\def\z{\vec{z}} %% open variables
\def\ep{\vec{o}}
\def\ga{\vec{p}}

@@

We define the vocabulary $\L = \{ S_1,\dots,S_t, \leq \}$, where $S_1,\dots,S_t$ have arity $b_1,\dots,b_t$. Let
\begin{eqnarray*}
\Truc &=& \{\A \mid \A \text{ is an } \L \text{-structure with a finite domain } A \text{ such that} \\
&& \leq \text{ is interpreted as a total order for } A \}.
\end{eqnarray*}
We also define a set of second order variables ${\cal X} = \{ X_i \mid i\in\N \}$ where $X_i$ has arity $a_i$, and for every $n \in \N$ there are infinite variables in ${\cal X}$ of arity $n$. A quantifier-free $\L$-formula is defined by the following grammar:
\begin{eqnarray*}
\varphi &::=& x = y \ \mid \ S_i(x_1,\dots,x_{b_i}), i \in \{1,\dots,t\} \ \mid \ x \leq y \ \mid \\
&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \\ 
&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi),
\end{eqnarray*}
where $x,y$ and $x_i$ are first order variables for every $i$. We now define an extended quantifier-free $\L$-formula as follows:
\begin{eqnarray*}
\varphi &::=& \alpha, \alpha \text{ is an FO-formula over } \L  \ \mid \\
&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \ \\
&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi).
\end{eqnarray*}
Let $\Y = (Y_1,\dots,Y_q)$ be a tuple of second-order variables of arity $c_1,\ldots,c_q$, and let $\y$ be a tuple of first order variables. For every $\L$-formula $\psi(\Y,\y)$, we define the function $f_{\psi(\Y,\y)}:\Truc \to \N$ as follows:
\begin{eqnarray*}
f_{\psi(\Y,\y)}(\A) &=& \mid \{ \langle \P, \e \rangle \mid \A \models \psi(\P,\e) \} \mid,
\end{eqnarray*}
for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_q)$ is a tuple of predicates of arity $c_1,\ldots,c_q$, and $P_i \subseteq A^{c_i}$ for every $i \in \{1,\ldots,q\}$, and $\e$ is a tuple of elements in $\A$.

Let $\Y = (Y_1,\dots,Y_r)$, and let $\y$ be a tuple of variables. A function $f:\Truc \to \N$ is in $\E1$ if there exists an extended quantifier-free $\L$-formula $\varphi(\x,\Y,\y)$ such that $f = f_{\exists \x \: \varphi(\x,\Y,\y)}.$\\

@@

For a given function $f$, we define $f \dotminus 1$ as follows:
\begin{eqnarray*}
f \dotminus 1(\A) =
\begin{cases}
f(\A)-1, & \text{if }f(\A) > 0 \\
0, & \text{if }f(\A) = 0.
\end{cases}
\end{eqnarray*}
for every $\L$-structure $\A \in \Truc$. A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$, it holds that $f \dotminus 1 \in \F$.

@@

\begin{theo}
$\E1$ is closed under substraction by one.
\end{theo}
\begin{proof}
Consider an $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$. Without loss of generality, let $\z = (z_1,\dots,z_d)$ and let $\X = (X_1,\dots,X_r)$. There are three possibilities regarding the size of the tuples of free variables $\z$ and $\X$: (1) $d>0$ and $r=0$ (2) $d=0$ and $r>0$ (3) $d,r>0$. This separates the proof in three cases:
\begin{enumerate}

@@

\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,\dots,x_d)$ and $\X = (X_1,\dots,X_r)$. That is,
\begin{eqnarray*}
f(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \exists \x \ \varphi(\x,\P) \} \mid \label{f1},
\end{eqnarray*}
for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_r)$ and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$. For the time being, suppose that
\begin{eqnarray}
\varphi(\x,\X) &=& \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)
\end{eqnarray}
where $n$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, $\y$ is a $p$-tuple of variables in $\x$, $\varphi^{-}(\X,\y)$ is a conjunction of negated predicates in $\X$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\beta(\x)$ is an FO-formula over $\L$ which mentions all variables in $\x$. Note that $\theta(\x)$ also mentions all variables in $\x$. We also assume that $(\x_1,\dots,\x_n,\y) = \x$. As an example, the following formula is of this form:
\begin{multline*}
\varphi(\x,\X) =  X_1(x_1,x_2) \wedge X_3(x_3) \wedge X_2(x_4,x_5) \wedge X_3(x_6) \wedge \neg X_1(x_7,x_8) \wedge \\ (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 \wedge x_2 = x_8 \wedge x_2 = x_5 \wedge x_8 < x_6 \wedge x_6 = x_7 ) \wedge \\ \forall z\big( S_1(x_1,z,x_2) \wedge x_3 = x_3 \wedge x_4 = x_4 \wedge x_5 = x_5 \wedge x_6 = x_6 \wedge x_7 = x_7 \wedge x_8 = x_8 \big),
\end{multline*}
where $\x = (x_1,x_2,x_3,x_4,x_5,x_6,x_7,x_8)$ and $\X = (X_1,X_2,X_3)$. Here, $n = 4$, $\lambda(1) = 1$, $\lambda(2) = \lambda(4) = 3$ and $\lambda(3) = 2$, $\x_1 = (x_1,x_2)$, $\x_2 = (x_3)$, $\x_3 = (x_4,x_5)$, $\x_4 = (x_6)$ and $\y = (x_7,x_8)$. Moreover, $\varphi^{-}(\X,\y) = \neg X_1(x_7,x_8)$, $\theta(\x) = (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 \wedge x_2 = x_8 \wedge x_2 = x_5 \wedge x_8 < x_6 \wedge x_6 = x_7 )$, which defines a total order on the partition of $\x$ $\{\{x_1,x_3,x_4\},\{x_2,x_5,x_8\},\{x_6,x_7\}\}$, and $\beta(\x) = \forall z\big( S_1(x_1,z,x_2) \wedge x_3 = x_3 \wedge x_4 = x_4 \wedge x_5 = x_5 \wedge x_6 = x_6 \wedge x_7 = x_7 \wedge x_8 = x_8 \big)$.

@@

\varphi^\prime(\x,\X) = \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)\wedge \\ \exists\u_1\cdots\exists\u_n\bigg[\alpha_{\min}(\u_1,\dots,\u_n) \wedge \bigg(\bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\u_i) \bigg) \vee \bigvee_{i=1}^r \exists \v\Big( X_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big) \bigg) \bigg] \label{f2}.

@@

\begin{lemma}  \label{first}

@@

\item[(a)] By contradiction, suppose that $\A\not\models\exists\x\,\varphi(\x,\P)$. That is, there is no assignment $\s$ to $\x$ such that $\A\models\varphi(\s,\P).$ Since $\d$ is such that $\A\models \bigwedge_{i=1}^n P_{\lambda(i)}(\d_i)$ and $\A\models \alpha(\d)$, it follows that $\A\not\models\varphi^{-}(\P,\ep)$ (since $\alpha(\d)=\theta(\d)\wedge\beta(\d)$). Therefore, there is an $i\in\{1,\ldots,n\}$ such that $\neg P_{\lambda(i)}(\d_i)$ appears in $\varphi^{-}(\P,\ep)$. Then, there is an $a_{\lambda(i)}$-tuple $\z$ in $\y$ such that $\neg X_{\lambda(i)}(\z)$ appears in $\varphi^{-}(\X,\x)$. We know that either (1) $\theta(\x)\models \z = \x_i$, (2) $\theta(\x)\models \varphi_{<,a_i}(\z,\x_i)$ or (3) $\theta(\x)\models \varphi_{<,a_i}(\x_i,\z)$. Considering that (2) and (3) are not possible given that both $\z$ and $\x_i$ are assigned the value $\d_i$ and $\A\models\theta(\d)$, we have that $\theta(\x)\models \z = \x_i$. But if this is the case, then $X_{\lambda(i)}(\x_i), \neg X_{\lambda(i)}(\z)$ and $\z = \x_i$ are all logical consequences of $\varphi(\x,\X)$, from which we conclude that $\A\not\models\exists\x\,\varphi(\x,\R^\prime)$ for every possible assignment $\R^\prime$ to $\X$, which contradicts the initial assumption that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$.
\item[(b)] Note that if $\A \models \alpha_{\min}(\c_1,\ldots,\c_n)$, then necessarily $\c_i = \d_i$ for $i\in\{1,\ldots,n\}.$ However, by the construction of $\P$, we see that 
$$\A\not\models\bigvee_{i = 1}^{n}\neg P_{\lambda(i)}(\d_i) \text{ and that } \A\not\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \d_j\Big).$$ Then, $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$.
\item[(c)] By contradiction, let $\P^\prime \neq \P$ be such that $\A\models\exists\x\,\varphi(\x,\P^\prime)$ and $\A\not\models\exists\x\,\varphi^\prime(\x,\P^\prime)$. We consider two cases: first, suppose that $\P^\prime$ is missing a tuple of $\P$. Let $i\in\{1,\ldots,n\}$ such that in $\d_i$ is not in $P^\prime_i$. Then $\A\models \neg P^\prime_i(\d_i)$, and so,$$\A\models \bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\d_i) \bigg), $$ from which we conclude that $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$. Second, suppose there is some predicate $\P^\prime_i$ in $\P^\prime$ which has a tuple that $P_i$ does not have. If this is the case, then $$\A\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big),$$ so $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$. On both cases, we have a contradiction.

@@

With this, we conclude that for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0,$ we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$

Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\R)$. Let $\P$ be an arbitrary assignment to $\X$. Since $\A\not\models\exists\x\,\varphi(\x,\P)$, we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\psi(\x,\P))$ for any formula $\psi(\x,\P)$. It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime(\x,\R)$. And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ it holds that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1,$ which was to be shown.
\end{proof}

We now continue with the general case, in which $\varphi(\x,\X)$ is an arbirtary extended quantifier-free $\L$-formula. By using a standard DNF transformation algorithm and considering FO-formulas over $\L$ as literals, we can find formulas $\gamma_i(\x,\X)$ with $i\in\{1,\ldots,\ell\}$ such that
\begin{eqnarray*}
\varphi(\x,\X) &\equiv& \gamma_1(\x,\X) \vee \gamma_2(\x,\X) \vee \dots  \vee \gamma_{\ell}(\x,\X),
\end{eqnarray*}
where, for every $i\in\{1,\ldots,\ell\}$, 
\begin{eqnarray*}
\gamma_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \gamma^{-}_i(\X,\y_i)  \wedge \delta_i(\x).
\end{eqnarray*}
The function $\lambda_i$ is defined analogously to the first part of the proof, the tuple $\x_{i,j}$ has $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, $\y_i$ has $p_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x.$ The formulas $\gamma^{-}_i(\X,\y_i)$, $\delta_i(\x)$, are also defined analogously to the first part of the proof (refer to \eqref{f1}). Let $g$ be a function that counts the number of possible orders over partitions on a $d$-tuple of variables. Let the formulas $\theta^i(\x)$ for $i\in\{1,\ldots,g(d)\}$ represent each of these orders over $\x$. Note that for every formula $\eta(\x)$,
\begin{eqnarray*}
\exists\x\:\eta(\x) &\equiv& \exists\x(\eta(\x)\wedge\theta^1(\x)) \vee \cdots \vee \exists\x(\eta(\x)\wedge\theta^{g(d)}(\x)).
\end{eqnarray*}
We define the following formulas $\xi_i(\x,\X)$, for $i\in\{1,\ldots,m\},$ as follows: 
\begin{eqnarray*}
\xi_1(\x,\X) &=& \gamma_1(\x,\X) \wedge \theta^1(\x), \\
& \vdots & \\
\xi_{g(d)}(\x,\X) &=& \gamma_1(\x,\X) \wedge \theta^{g(d)}(\x), \\
\xi_{g(d)+1}(\x,\X) &=& \gamma_2(\x,\X) \wedge \theta^1(\x), \\
& \vdots & \\
\xi_{2\cdot g(d)}(\x,\X) &=& \gamma_2(\x,\X) \wedge \theta^{g(d)}(\x), \\
& \vdots & \\
\xi_{(\ell-1) g(d)+1}(\x,\X) &=& \gamma_{\ell}(\x,\X) \wedge \theta^1(\x), \\
& \vdots & \\
\xi_{\ell \cdot g(d)}(\x,\X) &=& \gamma_{\ell}(\x,\X) \wedge \theta^{g(d)}(\x).
\end{eqnarray*}
Note that $m = \ell \cdot g(d).$ Having every disjunct with a total order allows us to eliminate the ones that are unsatisfiable for every $\L$-structure $\A$, that is, each $i\in\{1,\ldots,m\}$ such that for every assignment $\s$ to $\x,$ $f_{\xi_i(\s,\X)}(\A) = 0,$ for every $\A\in\Truc.$ Let $k$ be the number of disjuncts that are left after eliminating the unsatisfiable disjuncts. We use a function $\rho:\{1,\ldots,k\}\to\{1,\ldots,m\}$ such that $\varphi_i(\x,\X) = \xi_{\rho(i)}(\x,\X)$ is satisfiable, for every $i\in\{1,\ldots,k\}.$

\begin{claim}
For every assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$, then $f_{\varphi_i(\s,\X)}(\A) > 0.$
\end{claim}
\begin{proof}
By contradiction, assume that there is an assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$, and $f_{\varphi_i(\s,\X)}(\A) = 0.$ That is, there is no assignment $\R$ to $\X$ such that $\A\models\varphi_i(\s,\R).$ Since $\A\models\alpha_i(\s)$, it holds that $\A\models\theta_i(\s)$. Therefore, as we showed in part a) of Lemma \ref{first}, there is no assignment $\s^\prime$ to $\x$ such that $f_{\varphi_i(\s,\X)}(\A) > 0.$ This means $\varphi_i(\x,\X)$ is unsatisfiable, but, as we mentioned previously, all of such disjuncts have been eliminated, which is a contradiction.
\end{proof}
